---
title: "Analyze data"
date: "Compiled at `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'UTC')` UTC"
output: github_document
params:
  name: "02-analyze" # change if you rename file
---

```{r here, message=FALSE}
here::i_am(paste0(params$name, ".Rmd"), uuid = "a4069103-4402-4559-ba03-cca3df086442")
```

The purpose of this document is to create some state-based maps that show the current trajectory of COVID-19 cases. 
There will be two maps:

- seven-day average of newly-reported cases
- change in newly-reported cases vs. previous seven days

```{r packages}
library("conflicted")
library("readr")
library("dplyr")
library("albersusa")
library("ggplot2")
library("glue")

conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
```

```{r directories}
# create or *empty* the target directory, used to write this file's data: 
projthis::proj_create_dir_target(params$name, clean = TRUE)

# function to get path to target directory: path_target("sample.csv")
path_target <- projthis::proj_path_target(params$name)

# function to get path to previous data: path_source("00-import", "sample.csv")
path_source <- projthis::proj_path_source(params$name)
```

## Read data

```{r population}
population <- 
  read_csv(
    path_source("01-clean", "population.csv"),
    col_types = cols(state = "c", population = "d")
  ) %>%
  print()
```


```{r covid}
covid <- 
  read_csv(
    path_source("01-clean", "covid.csv"),
    col_types = cols(
      date = "D", 
      state = "c", 
      fips = "c", 
      cases = "d", 
      deaths = "d"
    )
  ) %>%
  print()
```

## Wrangle data

```{r covid_seven_day}
growth <- function(x) {
  # week-over-week growth, normalized by previous week
  
  # use "+ 1" in denominator to avoid division by zero
  result <- (x - dplyr::lag(x, 7)) / (dplyr::lag(x, 7) + 1) 
  
  # three digits should be more than enough
  round(result, 3)
}

per_100k <- function(x, pop) {
  result <- x / (pop / 1e5)
  
  round(result, 3)
}

covid_week <-
  covid %>%
  left_join(population, by = "state") %>%
  group_by(state) %>%
  arrange(date) %>%
  transmute(
    date,
    state,
    cases_total = cases, 
    cases_total_per100k = per_100k(cases_total, population),
    cases_avg_week = (cases - lag(cases_total, 7)) / 7,
    cases_avg_week_per100k = per_100k(cases_avg_week, population),
    cases_week_growth = growth(cases_avg_week),
    deaths_total = deaths,
    deaths_total_per100k = per_100k(deaths_total, population),
    deaths_avg_week = (deaths - lag(deaths_total, 7)) / 7,
    deaths_avg_week_per100k = per_100k(deaths_avg_week, population),
    deaths_week_growth = growth(deaths_avg_week)
  ) %>%
  print()
```

It might also be useful to have files for the most-recent day, each for cases and deaths.

```{r most-recent-cases}
covid_recent_cases <- 
  covid_week %>%
  filter(date == max(date)) %>%
  select(date, state, starts_with("cases")) %>%
  arrange(desc(cases_avg_week_per100k)) %>%
  print()
```

```{r most-recent-deaths}
covid_recent_deaths <- 
  covid_week %>%
  filter(date == max(date)) %>%
  select(date, state, starts_with("deaths")) %>%
  arrange(desc(deaths_avg_week_per100k)) %>%
  print()
```

## Plot data

Let's make some choropleth maps using [ggplot2](https://ggplot2.tidyverse.org/).

```{r}
map_recent_cases <- 
  usa_sf("laea") %>%
  left_join(covid_recent_cases, by = c(name = "state"))

date <- max(covid_recent_cases$date)

gg_cases <-
  ggplot(map_recent_cases, aes(fill = cases_avg_week_per100k)) +
  geom_sf(size = 0.25, color = "white") + 
  scale_fill_distiller(
    palette = "Oranges", 
    direction = 1,
    limits = c(0, NA)
  ) +
  labs(
    title = glue("Newly-reported COVID-19 cases, seven-day average as of {date}"),
    subtitle = "Data from the New York Times",
    fill = "cases\nper 100k"
  ) +
  theme_void() + 
  theme(
    legend.text.align = 1 # right-justify
  )

gg_cases
```
```{r}
# see https://github.com/tidyverse/ggplot2/issues/3738#issuecomment-583750802
mid_rescaler <- function(mid = 0) {
  function(x, to = c(0, 1), from = range(x, na.rm = TRUE)) {
    scales::rescale_mid(x, to, from, mid)
  }
}

gg_change <-
  ggplot(map_recent_cases, aes(fill = cases_week_growth)) +
  geom_sf(size = 0.25, color = "white") + 
  scale_fill_distiller(
    palette = "PuOr", 
    rescaler = mid_rescaler(),
    labels = scales::label_percent()
  ) +
  labs(
    title = glue("Week-over-week change in reported COVID-19 cases, as of {date}"),
    subtitle = "Data from the New York Times",
    fill = "change"
  ) +
  theme_void() +
  theme(
    legend.text.align = 1 # right-justify
  )

gg_change
```

## Write data

```{r write}
write_csv(covid_week, path_target("covid_week.csv"))
write_csv(covid_recent_cases, path_target("covid_recent_cases.csv"))
write_csv(covid_recent_deaths, path_target("covid_recent_deaths.csv"))

ggsave(path_target("cases.png"), plot = gg_cases, width = 7, height = 5)
ggsave(path_target("change.png"), plot = gg_change, width = 7, height = 5)
```

## Files written

These files have been written to ```r paste0("data/", params$name)```:

```{r list-files-target}
projthis::proj_dir_info(path_target())
```
