---
title: "Publish"
date: "Compiled at `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'UTC')` UTC"
output: github_document
params:
  name: "99-publish" # change if you rename file
---

```{r setup}
here::i_am(paste0(params$name, ".Rmd"), uuid = "ec845588-783a-4d74-9389-81c54875c3c3")

library("conflicted")
library("projthis")
library("here")
library("fs")
library("rprojroot")
```

The purpose of this document is to pick-and-choose from the data-files written earlier in this workflow in order to make them more-widely available.
In other circumstances, this may involve uploading data to an external service. Here, we will:

- put CSV files into a directory where we will feel confident they can be found going forward.
- put PNG files into a directory where the parent-project's README file can expect to find them.

```{r directories}
# create target directory to write *this* file's data: 
#  - all data written by this file should be written here
proj_create_dir_target(params$name)

# create accessor functions for data directories:
#  - get path to target directory: path_target("sample.csv")
#  - get path to previous data: path_data("00-import", "sample.csv")
path_target <- proj_path_target(params$name)
path_data <- proj_path_data(params$name)
```

## Transfer files

```{r csv-files}
file_copy(
  path_data("02-analyze", "covid_recent_cases.csv"), 
  path_target(),
  overwrite = TRUE
)

file_copy(
  path_data("02-analyze", "covid_recent_deaths.csv"), 
  path_target(), 
  overwrite = TRUE
)

file_copy(
  path_data("02-analyze", "covid_week.csv"), 
  path_target(), 
  overwrite = TRUE
)
```


```{r png-files}
dir_figs <- find_package_root_file("man", "figures") 
# create a directory in the package root

dir_create(dir_figs)

file_copy(path_data("02-analyze", "cases.png"), dir_figs, overwrite = TRUE)
file_copy(path_data("02-analyze", "change.png"), dir_figs, overwrite = TRUE)
```

## Files written

These files have been written to ```r paste0("data/", params$name)```:

```{r list-files-target}
proj_dir_info(path_target())
```
