---
title: "Publish"
date: "Compiled at `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'UTC')` UTC"
output: github_document
params:
  name: "99-publish" # change if you rename file
---

```{r here, message=FALSE}
here::i_am(paste0(params$name, ".Rmd"), uuid = "ec845588-783a-4d74-9389-81c54875c3c3")
```

The purpose of this document is to pick-and-choose from the data-files written earlier in this workflow in order to make them more-widely available.
In other circumstances, this may involve uploading data to an external service. 

Here, we will:

- put CSV files into a directory where we will feel confident they can be found going forward.
- put PNG files into a directory where the parent-project's README file can expect to find them.

```{r packages}
library("conflicted")
```

```{r directories}
# create or *empty* the target directory, used to write this file's data: 
projthis::proj_create_dir_target(params$name)

# function to get path to target directory: path_target("sample.csv")
path_target <- projthis::proj_path_target(params$name)

# function to get path to previous data: path_source("00-import", "sample.csv")
path_source <- projthis::proj_path_source(params$name)
```

## Transfer files

We are going to copy some files in to the target directory; we can write a wee function to make things easier:

```{r transfer-function}
copy_target <- function(path) {
  fs::file_copy(path, path_target())
}
```

We use our function to copy CSV files:

```{r csv-files}
copy_target(path_source("02-analyze", "covid_recent_cases.csv"))
copy_target(path_source("02-analyze", "covid_recent_deaths.csv"))
copy_target(path_source("02-analyze", "covid_week.csv"))
```


We use our function to copy PNG files:

```{r png-files}
copy_target(path_source("02-analyze", "cases.png"))
copy_target(path_source("02-analyze", "change.png"))
```

## Files written

These files have been written to ```r paste0("data/", params$name)```:

```{r list-files-target}
projthis::proj_dir_info(path_target())
```
